Deterministic Scheduling of Periodic Messages in a Network

### 2017-07-04 article/article.tex
corrections finales

     The second stage of our approach is very similar to the following scheduling problem: 
     we are given a set of jobss and each job has a \emph{release time} and a \emph{deadline}. 
     The problem is to schedule all jobs on a single processor so that no two jobs are scheduled at the same time, a job is always shedulded after its release time and job is finished before its deadline. Let us call $n$ the number of jobs, the problem can be solved in time $O(n^2\log(n))$~\cite{simons1978fast} when all jobs have the same running time and it gives a solution with the earliest possible deadline. On the other hand if the running times are different the problem is $\NP$-complete~\cite{lenstra1977complexity}. 
     The  polynomial time algorithm  which solves this scheduling problem is similar to the Greedy algorithm presented in the previous section. However, when it fails because a job will finish after its deadline, it changes the schedule of the last messages to solve find a possible schedule for the problematic job. It does it in a way that it will not fail on the same job a second time except if there is no solution, hence the polynomiality of the algorithm.

     We reduce our problem of setting the offsets of the backwards routes once the order of the forward routes is fixed to this
     scheduling problem. The backward routes are the jobs, the size of a message is the running time of a job,
     the deadline of a route is the deadline of the corresponding job and the smallest time at which it is eligible is the release time. Let us call {\bf Minimal Latency Scheduling (MLS)} this algorithm.
     Remark that we do not deal with the periodicity. A solution ${\cal M}$ the algorithm finds always respect $MT({\cal M}) < T_{max}$. On the other hand, it is guaranteed to be a $(P,\tau)$ periodic assignment only if $m_n - m_1 \leq P -\tau$ where $m_n$ is the largest offset and $m_1$ the smallest one. The algorithm we use minimize $m_n - m_1 $, so it may work quite well in practice (as show in Section.\ref{sec:resultsPALL}).

     We now present a variant of the previous algorithm that we call {\bf Periodic Minimal Latency Scheduling (PMLS)}
     which take into account the periodicity. We fix arbitrarily a backward route so that its message is the first to go through the central arc at time $t$. Then we modify the deadline of each forward route to be the minimum of their previous deadline and $t + P$.  We execute this algorithm for every possible first backward route while no solution is found. Since we run the previous algorithm at most $n$ times, the complexity is in $O(n^3\log(n))$. Remark that it will always find more solution than MLS,
     because if MLS find a solution with first route $r$ and such that $m_n - m_1 \leq P -\tau$, then this solution will be found by PMLS when it will fix $r$ as the first route. Moreover when PMLS finds a solution, it is always a $(P,\tau)$ periodic assignment
     since we guarantee that all messages are scheduled in the same period of time of size $P$ and it satisfies $MT({\cal M}) < T_{max}$ by construction of the deadlines. Remark that it is possible that there is a solution and PMLS do not find it  
     since we do not allow to chose an offset for a route after $t+P- \tau$ which modulo $P$ does not collide with another route.

@@

    First we try to understand what is the best choice of order for the first phase of the algorithm.
    We fix the number of routes to $8$, and we  draw uniformly the weigths of the arcs between $0$ and $20000$ slots.
    For a given topology, we define the {\em margin} as the difference between $T_{max}$ and twice the longest route. 
    Hence, for a given topology giving the margin or $T_{max}$ is the same, but since we will consider many topologies 
    wich may have different longest route's size, it will be more meaningful to set the margin rather than $T_{max}$.
    In fact the margin represents the latency imposed by the communication process without taking into account the physical length 
    of the network which cannot be changed. In our experiments the margin range from  $0$ to $3000$ slots.
   We look at two different regimes: first, we take a period equal to $1,05n\tau$ to represent a very loaded network
   and then we chose $1,5n\tau$ for a mildly loaded network. Chosing a larger one is not relevant since then we know how to solve the problem without waiting times as shown in Section~\ref{sec:exp_PAZL}. We represent the success rate with regards to the margin, 
   for the five kind of orders using the Greedy algorithm for the second stage, computed over $10000$ random star topologies. Note that for the random order, we compute $100$ random orders and count it as a success as soon as there is a solution for one order.

@@

       \begin{center}
      \includegraphics[scale=0.4]{departs_gp_30000.pdf}
      \end{center}
      \caption{Success rate of differents orders with $P = 30000$}
      \label{fig:success30000}
     \end{figure}

     As expected, the  success rate is much better for large periods. According to our results, sending the messages from the sortest to the longest route or arc does not work well. It corresponds to the policy of Proposition~\ref{prop:SL} which does not work well when the routes are long as it is the case. Sending from the longest to the shortest route or arc works better and it seems that sorting the routes according to the length of $(c_t,t_i)$ is better at least in loaded network.
     FInally the random order is much better than the other orders, when the load is not high ($P = 30000$) there is always a solution with margin $0$.

    the two SL sending orders are not good, in any cases. However, the LS orders seem to increase when the flexibility is large enough, but in the two regimes, they appear to be stagnating to a lower success rate than Random. As we can see the random finds near to 100\% of the solutions when the period is large enough, even when we give no flexibility. 
     %On average, it is necessary to draw 6 randoms sending order to find a solution, when the period is large, but in some particularly hard instances, it can be impossible to find a solution, even with 100 randoms sending orders.

     To observe  the impact of the number of random sending orders drawn, we compared the success rate of the following number of drawing, on graphs in which the links are drawn  between $0$ and $20000$ slots, with a period $P=20000$, and a flexibility of $0$. The following table present the average success rate of each number of sending orders, on $1000$ instances.
       \begin{figure}[H]

@@

     In our case of $8$ routes, the number of different possible sending orders is $8! \simeq 40 000$. By drawing $1000$ random sending orders, we explore near to $2,5 \%$ of the cases, which seems enough to find as many solutions as if generate more random sending orders.

     We now want to compare the performances of the different algorithms on the way back. Since GP showed some excellent results on large periods (see Fig~\ref{fig:success30000}), it is more interesting to focus on the behavior of the algorithms on short periods. The following experience is realized on graph with $8$ routes (links between $0$ and $20000$), and with a period of $21000$. Since we just conclude that drawing $1000$ sending order is enough, with $8$ routes, we now generate $1000$ randoms sending orders instead of $100$.

@@

     The algorithm MS, that does not consider the periodicity for the scheduling is far worst than the others, and, even if GP seems to converge too $100\%$ of success rate when the flexibility is large enough, MSP is always strictly better. We remark that when the flexibility is low ($<500$) there is still near to $5\%$ of failure for MSP.

   Thus, using the Random sending order combined with MSP for the way back seems to be a very good algorithm to solve PALL. Nevertheless, there is still a few number of instances in the worst area of parameters (very short period, and very low flexibility), that are impossible to solve with our algorithm.

 To evaluate the performance of our best algorithm to find solutions, that is, generating $1000$ randoms sending orders and using MSP on the way back, we compared it's performance against the actual way to manage the messages in a network. Nowadays, the networks are managed through a stochastic model. The messages in the nodes are assigned following a FIFO rule. Even if this policy works well, in practice when the network is not too loaded, it does not ensure a minimum latency. Thus, some latency critical applications like C-RAN needs some deterministic algorithms. To model such a network, we generate some randoms offsets in the sources. Then we simulate the behavior of a stochastic network during a great number of periods.

We simulated a stochastic network on $1000$ instances with $8$ long routes (links between $0$ and $20000$). The results are not really different if we use short routes. The flexibility needed by an instance of the network is computed by taking the message which have the highest process time, at which we subtract the weight of the longest route. We measure, for periods from $20000$ ( network fully loaded), to $50000$ ($40 \%$ of load), the medians, the first, and the third quartiles of $1000$ results.

@@

       \begin{center}
     % \includegraphics[scale=0.4]{stochastic.pdf}
       \input{figures/stochastic}
      \end{center}
      \caption{Needed flexibility on stochastic networks according to the load.}
      \label{fig:sto}
     \end{figure}

    We also made a linear search to find the minimum flexibility needed by our algorithm on those loads to find a solution. There are few cases, when both the flexibility and the period are too restricted in which there is no solutions: with a period of $20000$, our algorithm needs in average $\simeq 80$ slots of flexibility to find a solution. Otherwise, the algorithm finds a solution with $0$ flexibility for every periods $>20000$. On the other hand, the stochastic policy does not ensure a minimal latency. As we can see in Fig~\ref{fig:sto}), the more the network is loaded, the higher the latency is. Even if the first quartile is to $0$ when the network is not highly loaded, many instances needs some flexibility. Moreover, the latency is not guaranteed with that policy. Indeed, even if half of the cases of our simulations works with less than $1000$ slots of flexibility on a light loaded network, that could be reasonable for specific applications, $25\%$ of the cases needs more than $2000$ slots of flexibility. With our algorithm, we control the latency, and here, we have optimal solutions that makes it really appropriate.

@@

   We plan to generalize our study of the PALL problem to more real topologies,
   such as trees, cycles or bounded treewidth graphs. 
   We would like to design FPT algorithms in the number of routes for PAZL and PALL on the star topology, eventually by relaxing the constraint on the maximal process time into an constraint on the average process time, which seems much more tractable.

   We could also study variation of our problem. Instead of minimizing the maximum process time, we could try to minimize
   the sum of the waiting time, a linear objective which could make linear programming useful.
   Moreoever we could allow preemption, that is the messages are allowed to be cut into pieces, which would certainly change the complexity of the problem. 
   Finally, the routes may not be fixed but chosen in the graph to minimize the maximum process time, which would make the problem even more difficult (maybe $\Pi_2$-complete instead of $\NP$-complete).

>>>>

     The problem to solve in the second stage of our approach is very similar to the following scheduling problem: 
     we are given a set of jobs and each job has a \emph{release time} and a \emph{deadline}. 
     The problem is to schedule all jobs on a single processor, that is choosing the time at which they are computed, so that no two jobs are scheduled at the same time. A job is always scheduled after its release time and it must be finished before its deadline. Let us call $n$ the number of jobs, the problem can be solved in time $O(n^2\log(n))$~\cite{simons1978fast} when all jobs have the same running time and it gives a solution with the earliest possible deadline. On the other hand if the running times are different the problem is $\NP$-complete~\cite{lenstra1977complexity}. 
     The  polynomial time algorithm  which solves this scheduling problem is similar to the Greedy algorithm presented in the previous section. However, when it fails because a job will finish after its deadline, it changes the schedule of the last messages to find a possible schedule for the problematic job. The change in the scheduling is so that the algorithm cannot fail on the same job a second time except if there is no solution, hence the polynomiality of the algorithm.

     We reduce our problem of setting the offsets of the backwards routes once the order of the forward routes is fixed to this
     scheduling problem. The backward routes are the jobs, the size of a message is the running time of a job,
     the deadline of a route is the deadline of the corresponding job and the smallest time at which it is eligible is the release time. Let us call {\bf Minimal Latency Scheduling (MLS)} this algorithm.
     Remark that we do not deal with the periodicity. When MLS finds a solution ${\cal M}$ its always satisfies $MT({\cal M}) < T_{max}$. On the other hand, it is a $(P,\tau)$ periodic assignment only if $m_n - m_1 \leq P -\tau$ where $m_n$ is the largest offset and $m_1$ the smallest one. The algorithm we use minimize $m_n - m_1 $, so it may work quite well in practice (as shown in Section.\ref{sec:resultsPALL}).

     We now present a variant of the previous algorithm that we call {\bf Periodic Minimal Latency Scheduling (PMLS)}
     which take into account the periodicity. We fix arbitrarily a backward route so that its message is the first to go through the central arc at time $t$. Then we modify the deadline of each forward route to be the minimum of their previous deadline and $t + P$.  We execute this algorithm for every possible first backward route while no solution is found. Since we run the previous algorithm at most $n$ times, the complexity is in $O(n^3\log(n))$. Remark that it will always find more solution than MLS,
     because if MLS find a solution with first route $r$ and such that $m_n - m_1 \leq P -\tau$, then this solution will be found by PMLS when it selects $r$ as the first route. Moreover when PMLS finds a solution, it is always a $(P,\tau)$ periodic assignment
     since we guarantee that all messages are scheduled in a period of time of size $P$ and it satisfies $MT({\cal M}) < T_{max}$ by construction of the deadlines. Remark that it is possible that there is a solution and PMLS does not find it  
     since we do not allow to choose an offset for a route after $t+P- \tau$ which modulo $P$ may not collide with another route.

@@

    In this section, we first try to understand what is the best choice of order for the first stage of the algorithm.
    We fix the number of routes to $8$ to make comparisons with the results of Section~\ref{sec:exp_PAZL} easier. 
    We draw uniformly the weights of the arcs between $0$ and $20000$ slots.
    For a given topology, we define the {\em margin} as the difference between $T_{max}$ and twice the longest route. 
    Hence, for a given topology choosing the margin or $T_{max}$ is the same, but since we will consider many topologies 
    which may have different longest route's size, it will be more meaningful to set the margin rather than $T_{max}$.
    In fact the margin represents the latency imposed by the communication process without taking into account the physical length of the network which cannot be changed. In our experiments the margin range from  $0$ to $3000$ slots.
   We look at two different regimes: we take a period equal to $1,05n\tau$ to represent a very loaded network
   and $1,5n\tau$ for a mildly loaded network. Choosing a larger period is not relevant since then we know how to solve the problem even without waiting times as shown in Section~\ref{sec:exp_PAZL}. We represent the success rate with regards to the margin, for the five kind of orders using the Greedy algorithm for the second stage, computed over $10000$ random star topologies. Note that for the random order, we compute $1000$ random orders and count it as a success as soon as there is a solution for one order.

@@

       \begin{center}
      \includegraphics[scale=0.4]{departs_gp_30000.pdf}
      \end{center}
      \caption{Success rate of different orders with $P = 30000$}
      \label{fig:success30000}
     \end{figure}

     As expected, the  success rate is much better for large periods. According to our experiments, sending the messages from the shortest to the longest route or arc does not work well. It corresponds to the policy of Proposition~\ref{prop:SL} which does not work well when the routes are long as it is the case in these experiments. Sending from the longest to the shortest route or arc works better and it seems that sorting the routes according to the length of the last arc rather than the route is better, at least in a loaded network.

     The random order is is the best by far. When the load is not too high ($P = 30000$) there is always a solution with margin $0$.
     Note that, when disallowing waiting times, there were instance without solutions for these parameters (see Section~\ref{sec:exp_PAZL}), which justifies the interest of studying the PALL problem. Remark that the random order corresponds to the best of $1000$ orders drawn uniformly which means that a computation can be up to $1000$ time slower than for a fixed order. The choice of the number of random orders drawn yields a trade-off between the computation time and the success rate. Therefore we investigate the success rate of our algorithm with regards to the number of random orders drawn, a period of $21000$ and a margin $0$. The following table presents the average success rate for each number of sending orders, on $1000$ instances.
       \begin{figure}[H]

@@

    The number of different orders is $7!$ since we have $8$ routes and the solutions are invariant up to a circular permutation of the order. Therefore when doing $1000$ random draws, we explore less than $20\%$ of the possible orders but it is already enough to get the best results on average.

     We now want to compare the performances of the three different algorithms used in the second stage. Since GD showed excellent results on mild loads, it is more interesting to focus on the behavior of the algorithms on high load. The following experience is realized on graph with $8$ routes, weights of the arcs drawn between $0$ and $20000$ and with a period of $21000$. We choose the best possible first stage found in the previous experiments, that is trying $1000$ random orders.

@@

     The algorithm MLS performs poorly, worst than GP and PMLS, which shows that taking into account the periodicity is fundamental.
     The GP algorithms is close to $100\%$ success rate for margins larger than $1500$ while the PMLS finds a solution in more than $95\%$ of the experiments, even with a margin $0$. Therefore, for the worst possible constraints on load and margin, there are a few instances for which we do not find a solution. However, with a margin of $500$, which corresponds to about $0.025$ms of additional delay with the chosen parameters, we always find a solution.

     Now that we have found the best amongst the algorithms solving PALL, we need to compare its performances against the actual way to manage the messages in a network. The classical way to manage messages in the network is statistical multiplexing, 
     where there is a FIFO buffer in each node of the network to deal with the collisions. The time at which the messages are sent in the network is not computed as in our approach, thus we fix the offsets of each route to some random value.
     Even if this policy seems to work in practice when the network is not too loaded, it does not give any guarantee on the latency. Remark that the process is not periodic, therefore we must measure the process time of each route over several periods if we want to compute its maximum. We choose to simulate it for $1000$ periods and we have observed that it usually stabilizes after $10$ periods. The margin is also defined as the maximum process time, computed as explained, minus twice the size of the longest route.

     We draw $1000$ instances of a star topology with $8$ routes, with weights on the arcs drawn between $0$ and $20000$. Note that the results are almost the same when the arcs are small. We measure, for periods from $20000$ (network fully loaded), to $50000$ ($40 \%$ of load), the median, the first, and the third quartiles of the margins computed in $1000$ simulations using FIFO buffer to resolve collisions.

@@

       \begin{center}
     % \includegraphics[scale=0.4]{stochastic.pdf}
       \input{figures/stochastic}
      \end{center}
      \caption{Needed margin with respect to the period for a stochastic policy}
      \label{fig:sto}
     \end{figure}

     We also computed the minimum margin needed by PMLS to find a solution. With a period of $20000$, our it needs in average $\simeq 80$ slots of flexibility to find a solution. Otherwise, the algorithm finds a solution with a margin $0$ for every period larger than $20000$. On the other hand, the simulations show clearly that the stochastic policy does not ensure a minimal latency. As we can see in Fig~\ref{fig:sto}), the more the network is loaded, the higher the latency is. When the network is highly loaded,
     we have instances with a margin of about $10000$ which corresponds to half the period, that is $0.5$ms. 
     Even when the network is lightly loaded,  a quarter of the instances have a margin of more than $2000$
     which is $x$ times more than PMLS for a network fully loaded ! 
     We feel that it strongly justifies the use of a deterministic sending scheme for latency critical applications such as our C-RAN motivating problem.

@@

   We plan to generalize our study of the PALL problem to other topologies,
   such as trees, cycles or bounded treewidth graphs. 
   We would like to design FPT algorithms in the number of routes for PAZL and PALL on the star topology, eventually by relaxing the constraint on the maximal process time into an constraint on the average process time, which seems much more tractable.

   We could also study variation of our problem. Instead of minimizing the maximum process time, we could try to minimize
   the sum of the waiting time, a linear objective which could make linear programming useful.
   Moreover we could allow preemption, that is the messages are allowed to be cut into pieces, which would certainly change the complexity of the problem. 
   Finally, the routes may not be fixed but chosen in the graph to minimize the maximum process time, which would make the problem even more difficult (maybe $\Pi_2$-complete instead of $\NP$-complete).
